The guts of the dispRity package

### 2023-11-27 inst/gitbook/06_guts.Rmd
added MCMCglmm.variance and updates on randtest.dispRity
added example for get.tree in vignette

@@

>>>>

#### `name.subsets`

This function gets you the names of the subsets in a `dispRity` object as a vector.

```{r}
## What are they called?
name.subsets(disparity)
```

@@

Note that `get.tree` can also be used to extract trees from different subsets (custom or continuous/discrete subsets).

For example, if we have three time bins like in the example below we have three time bins and we can extract the subtrees for these three time bins in different ways using the option `subsets` and `to.root`:

```{r}
## Load the Beck & Lee 2014 data
data(BeckLee_tree) ; data(BeckLee_mat99) ; data(BeckLee_ages)

## Time binning (discrete method)
## Generate two discrete time bins from 120 to 40 Ma every 20 Ma
time_bins <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree,
                            method = "discrete", time = c(120, 100, 80, 60),
                            inc.nodes = TRUE, FADLAD = BeckLee_ages)

## Getting the subtrees all the way to the root
root_subsets <- get.tree(time_bins, subsets = TRUE)

## Plotting the bin contents
old_par <- par(mfrow = c(2,2))
plot(BeckLee_tree, main = "original tree", show.tip.label = FALSE)
axisPhylo()
abline(v = BeckLee_tree$root.time - c(120, 100, 80, 60))
for(i in 1:3) {
     plot(root_subsets[[i]], main = names(root_subsets)[i],
          show.tip.label = FALSE)
     axisPhylo()
}
par(old_par)
```

But we can also extract the subtrees containing only branch lengths for the actual bins using `to.root = FALSE`:

```{r}
## Getting the subtrees all the way to the root
bin_subsets <- get.tree(time_bins, subsets = TRUE, to.root = FALSE)

## Plotting the bin contents
old_par <- par(mfrow = c(2,2))
plot(BeckLee_tree, main = "original tree", show.tip.label = FALSE)
axisPhylo()
abline(v = BeckLee_tree$root.time - c(120, 100, 80, 60))
for(i in 1:3) {
     plot(bin_subsets[[i]], main = names(bin_subsets)[i],
          show.tip.label = FALSE)
     axisPhylo()
}
par(old_par)
```

This can be useful for example for calculating the branch lengths in each bin:

```{r}
## How many cumulated phylogenetic diversity in each bin?
lapply(bin_subsets, function(tree) sum(tree$edge.length))
```
